---
title: "Non-Seasonal Time Series Analysis"
author: 'Aston Glen Noronha (CWID: 20012232)'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<span style="font-size: 18px;font-weight: bold;">Introduction and Motivation </span>

<span style="font-size: 18px;">Stock market data can be interesting to analyze and as a further incentive, strong predictive models can have large financial payoff. The amount of financial data on the web is seemingly endless. A large and well-structured dataset on a wide array of companies can be hard to come by. Here I am using a dataset containing Twitter stock prices from 2013 to 2019.</span>

<span style="font-size: 18px;">The goal of this project is to predict and forecast daily return values for a particular stock. Given the highly 
volatile nature of stock data, we will fit a univariate GARCH model to achieve our goal of predicting the daily returns value to allow you to make statistically informed trades!</span>

<span style="font-size: 18px;font-weight: bold;">Data Description </span>

<span style="font-size: 18px;">Data Range : From November 2013 to December 2019 </span>

<span style="font-size: 18px;">Dataset Description : The dataset contains Twitter Stock Prices </span>

<span style="font-size: 18px;">Datasource Description : The data is from Kaggle website and can be accessed using the link: https://www.kaggle.com/datasets/maharshipandya/twitter-stocks-dataset</span>

<span style="font-size: 18px;">Date format : yyyy-mm-dd </span>

<span style="font-size: 18px;">Open : price of the stock at market open </span>

<span style="font-size: 18px;">High : Highest price reached in the day </span>

<span style="font-size: 18px;">Low : Lowest price reached in the day</span>

<span style="font-size: 18px;">Close : price of the stock at market close</span>

<span style="font-size: 18px;">Volume : Number of shares traded</span>

```{r}
library(stats)
library(tseries)
library(tidyverse)
library(TSA)
library(ggplot2)
library(rugarch)
library(forecast)
library(dplyr)
```
<span style="font-size: 18px;font-weight: bold;">Let's take a look at the dataframe we are working with</span>
```{r}
df <- read.csv("C:/Users/Aston/Downloads/Academic Projects/Time Series Project/twitter-stocks.csv")
df <- select(df, Date, Close)

df$Date <- as.Date(df$Date, "%Y-%m-%d")
num_rows <- nrow(df)
new_dates <- seq(from = as.Date("2013-11-01"), by = "days", length.out = num_rows)
df$Date <- new_dates
start_date <- as.Date("2013-11-01")
end_date <- as.Date("2019-12-31")
filtered_data <- subset(df, Date >= start_date & Date <= end_date)
head(filtered_data)
tail(filtered_data)
```

<span style="font-size: 18px;font-weight: bold;">Initial look at the data:</span>
```{r}
ggplot(filtered_data, aes(x = Date, y = Close)) +
  geom_line() +
  labs(title = "Time Series of Close Prices(Original)",
       x = "Date",
       y = "Close Price")
train_data <- subset(df, Date >= start_date & Date <= as.Date("2019-10-31"))
test_data <- subset(df, Date > as.Date("2019-10-31") & Date <= end_date)
data1 <- train_data$Close
```
```{r}
acf(data1, lag.max = 150, main='ACF of Original data')
```
```{r}
pacf(data1, lag.max = 150, main='PACF of Original data')
```

<span style="font-size: 18px;font-weight: bold;">EACF of the Original data</span>
```{r}
eacf(data1)
```
<span style="font-size: 18px;">From the ACF we see that it is not cutting off and constantly dying. From the PACF, it seems like AR(1) model. From the EACF, we can try different ARMA orders to try and capture all the dependencies.</span>

<span style="font-size: 18px;font-weight: bold;">Checking for stationarity</span>
```{r}
adf_test = adf.test(data1)
print(adf_test)
```
<span style="font-size: 18px;">Since p-value=0.3565 > 0.05, the TS is non-stationary</span>

<span style="font-size: 18px;font-weight: bold;">Differencing to make the data stationary</span>
```{r}
close_diff <- diff(data1)
plot(close_diff, type = 'l', xlab = 'Date', ylab = 'Value', main = 'Line graph for log-differencing')
```
```{r}
adf_test = adf.test(close_diff)
print(adf_test)
```
<span style="font-size: 18px;">Since p-value=0.01 < 0.05, the TS is stationary</span>
```{r}
acf(close_diff, main='ACF of log-differenced data')
pacf(close_diff, main='PACF of log-differenced data')
```

<span style="font-size: 18px;font-weight: bold;">EACF of the differenced data</span>
```{r}
eacf(close_diff)
```
<span style="font-size: 18px;font-weight: bold;">Functions for ARIMA models</span>
```{r}
plot_arima_resid <- function(model) {
  resid <- resid(model)
  resid_ts <- ts(resid, start=c(2013, 311), end=c(2020,300), frequency=365)
  p <- model$arma[1]
  q <- model$arma[2]
  plot(resid_ts, xlab = "Year", ylab = 'Residuals', main=paste('Residual Plot',"(", p, ", 1, ", q, ")"))
  acf(resid_ts[1:length(resid_ts)],lag.max=150, main=paste('ACF of Residual',"(", p, ", 1, ", q, ")"))
  pacf(resid_ts[1:length(resid_ts)],lag.max=150, main=paste('PACF of Residual',"(", p, ", 1, ", q, ")"))
}

my_df <- data.frame(model = character(),
                    AIC = numeric(),
                    BIC = numeric(),
                    Shapiro = round(numeric(),3),
                    Ljung = round(numeric(),3))
new_row <- data.frame(model='',AIC=0,BIC=0,Shapiro=0,Ljung=0)

plot_arima_residuals <- function(arima_model, my_df) {
  resid <- residuals(arima_model)
  resid_ts <- ts(resid, start=c(2013, 311), end=c(2020,300), frequency=365)
  p <- arima_model$arma[1]
  q <- arima_model$arma[2]
  order <- paste("(", p, ", 1, ", q, ")", sep = "")
  plot_arima_resid(arima_model)
  qqnorm(resid_ts, main=paste('Residuals plot',order))
  qqline(resid_ts)
  
  shap=shapiro.test(resid_ts)
  
  ljung=Box.test(resid, lag = 20, type = "Ljung-Box")
  
  cat("AIC:", AIC(arima_model),"\n")
  cat("BIC:", BIC(arima_model),"\n")
  model_name <- paste("ARIMA", paste(order, collapse = ','), sep = " ")
  if (!model_name %in% my_df$model) {
    new_row <- data.frame(model = model_name,
                          AIC = AIC(arima_model),
                          BIC = BIC(arima_model),
                          Shapiro = round(shap$p.value, 3),
                          Ljung = round(ljung$p.value, 3))
    
    my_df <- rbind(my_df, new_row)
  }
}
```
```{r}
arima011 <- arima(x=close_diff, order=c(0,1,1))
arima110 <- arima(x=close_diff, order=c(1,1,0))
arima111 <- arima(x=close_diff, order=c(1,1,1))
arima210 <- arima(x=close_diff, order=c(2,1,0))
arima211 <- arima(x=close_diff, order=c(2,1,1))
arima012 <- arima(x=close_diff, order=c(0,1,2))
arima112 <- arima(x=close_diff, order=c(1,1,2))
arima212 <- arima(x=close_diff, order=c(2,1,2))
arima310 <- arima(x=close_diff, order=c(3,1,0))
arima311 <- arima(x=close_diff, order=c(3,1,1))
arima312 <- arima(x=close_diff, order=c(3,1,2))
arima013 <- arima(x=close_diff, order=c(0,1,3))
arima113 <- arima(x=close_diff, order=c(1,1,3))
arima213 <- arima(x=close_diff, order=c(2,1,3))
arima313 <- arima(x=close_diff, order=c(3,1,3))
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(0,1,1)</span>
```{r}
my_df <- plot_arima_residuals(arima011, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(1,1,0)</span>
```{r}
my_df <- plot_arima_residuals(arima110, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(1,1,1)</span>
```{r}
my_df <- plot_arima_residuals(arima111, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(2,1,0)</span>
```{r}
my_df <- plot_arima_residuals(arima210, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(2,1,1)</span>
```{r}
my_df <- plot_arima_residuals(arima211, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(0,1,2)</span>
```{r}
my_df <- plot_arima_residuals(arima012, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(1,1,2)</span>
```{r}
my_df <- plot_arima_residuals(arima112, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA2,1,2)</span>
```{r}
my_df <- plot_arima_residuals(arima212, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(3,1,0)</span>
```{r}
my_df <- plot_arima_residuals(arima310, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(3,1,1)</span>
```{r}
my_df <- plot_arima_residuals(arima311, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(3,1,2)</span>
```{r}
my_df <- plot_arima_residuals(arima312, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(0,1,3)</span>
```{r}
my_df <- plot_arima_residuals(arima013, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(1,1,3)</span>
```{r}
my_df <- plot_arima_residuals(arima113, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(2,1,3)</span>
```{r}
my_df <- plot_arima_residuals(arima213, my_df)
```
<span style="font-size: 18px;font-weight: bold;">ARIMA(3,1,3)</span>
```{r}
my_df <- plot_arima_residuals(arima313, my_df)
```
<span style="font-size: 18px;font-weight: bold;">BIC Values of the models</span>
```{r}
BIC_values <- c(BIC(arima011), BIC(arima110), BIC(arima111),
                BIC(arima210), BIC(arima211), BIC(arima012), BIC(arima112), BIC(arima212),
                BIC(arima310), BIC(arima311), BIC(arima312), BIC(arima013), BIC(arima113), BIC(arima213), BIC(arima313))
BIC_values
```
<span style="font-size: 18px;font-weight: bold;">Lowest BIC Value</span>
```{r}
min_BIC <- min(BIC_values)
min_BIC
```
```{r}
min_BIC_index <- which.min(BIC_values)
min_BIC_index
```
<span style="font-size: 18px;">The lowest BIC values is from the model ARIMA(0,1,1). Therefore, we continue with the residual analysis to determine if the model has captured all the dependencies.</span>

<span style="font-size: 18px;font-weight: bold;">Residual Analysis for ARIMA(0,1,1)</span>

```{r}
residuals <- residuals(arima011)
acf(residuals, main = 'ACF of Residuals')
pacf(residuals, main = 'ACF of Residuals')
```

<span style="font-size: 18px;font-weight: bold;">Ljung's Box Test</span>
```{r}
tsdiag(arima011)
```

<span style="font-size: 18px;">From the above plot, we can see that most of the p-values are above the confidence interval, so lets try and forecast for the next 60 days.</span>
```{r}
predicted_values <- predict(arima011, n.ahead = 60)
original_forecasts <- diffinv(predicted_values$pred, lag = 1, differences = 1, xi = 39.84)
original_forecasts <- original_forecasts[original_forecasts != 0]
original_forecasts
```
```{r}
plot(test_data$Date, test_data$Close, main = 'Original Data vs Forecast for 2019', type = 'l', col = 'blue', xlab = 'Date', ylab = 'Employees')

lines(test_data$Date, original_forecasts, col = 'red', type = 'l')

legend('topleft', legend = c('Original Data', 'Forecasts'), col = c('blue', 'red'), lty = 1)
```

<span style="font-size: 18px;">As we can see from the forecast against the original data, we can say that the ARIMA model has not captured the dependencies. The reason this happens is because the variance of the original data is not constant. We go ahead and implement GARCH model to our data. </span>


<span style="font-size: 18px;font-weight: bold;">GARCH Model</span>
```{r}
return = diff(log(train_data$Close))^2

#Try with absolute and square of the returns
acf(return, lag.max = 100, main='ACF of return')
pacf(return, lag.max = 100, main='PACF of return')
```
```{r}
garch(x=return, grad='numerical', trace=FALSE)
garch_spec <- ugarchspec(variance.model = list(garchOrder = c(2,2)),
                         mean.model = list(armaOrder = c(2,2)),
                         distribution.model = 'std')

garch_fit <- ugarchfit(spec = garch_spec, data = return, solver = 'hybrid')
summary(garch_fit)
garch_fit
```
<span style="font-size: 18px;font-weight: bold;">Residual Analysis</span>
```{r}
resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')
```

<span style="font-size: 18px;font-weight: bold;">Forecasting and Simulating</span>
```{r}
sfinal <- garch_spec
setfixed(sfinal) <- as.list(coef(garch_fit))
coef(garch_fit)
f2019 <- ugarchforecast(data = return, fitORspec = sfinal, n.ahead = 60)
plot(sigma(f2019))
sim <- ugarchpath(spec = sfinal, m.sim = 2, n.sim = 1*60, rseed = 123)
p <- 39.84*apply(fitted(sim), 2, 'cumsum') + 39.84
matplot(p, type = "l", lwd = 3)
```

<span style="font-size: 18px;font-weight: bold;">gjrGARCH Model</span>
```{r}
garch_spec <- ugarchspec(variance.model = list(model = 'gjrGARCH'),
                         mean.model = list(armaOrder = c(2,2)),
                         distribution.model = 'std')

garch_fit <- ugarchfit(spec = garch_spec, data = return, solver = 'hybrid')
summary(garch_fit)
garch_fit

resid_norm = residuals(garch_fit)
qqnorm(resid_norm, main='QQ plot of residuals from gjr-GARCH ')
qqline(resid_norm)
acf(resid_norm, lag.max = 60, main='ACF of residual')
pacf(resid_norm, lag.max = 60, main='PACF of residual')

sfinal <- garch_spec
setfixed(sfinal) <- as.list(coef(garch_fit))
coef(garch_fit)
f2019 <- ugarchforecast(data = return, fitORspec = sfinal, n.ahead = 60)
sim <- ugarchpath(spec = sfinal, m.sim = 2, n.sim = 1*60, rseed = 123)
p <- 39.84*apply(fitted(sim), 2, 'cumsum') + 39.84
matplot(p, type = "l", lwd = 3)

```

